---
title: "Q1 Saratoga"
author: "Simran, Swatee & David"
date: "06/03/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(mosaic)
library(foreach)
library(knitr)
library(rmarkdown)
library(FNN)
library(dplyr)
data(SaratogaHouses)
summary(SaratogaHouses)
load("/Users/msimran28/Q1 Saratoga.R")
```

#Variables defined
```{r chunk1, echo = FALSE, include=FALSE}
SaratogaHouses$new <- ifelse(SaratogaHouses$age == 0, 1,0)
#SaratogaHouses$new

housevalue <- SaratogaHouses$price - SaratogaHouses$landValue
# head(housevalue)
heating_electric <- SaratogaHouses[grep("electric", SaratogaHouses$heating), ]
# head(heating_electric)
heating_steam <- SaratogaHouses[grep("hot water/steam", SaratogaHouses$heating), ]
# head(heating_steam)
heating_hotair <- SaratogaHouses[grep("hot air", SaratogaHouses$heating), ]
# head(heating_hotair)

fuel_oil <- SaratogaHouses[grep("oil", SaratogaHouses$fuel), ]
# head(fuel_oil)
fuel_gas <- SaratogaHouses[grep("gas", SaratogaHouses$fuel), ]
# head(fuel_gas)
fuel_electric <- SaratogaHouses[grep("electric", SaratogaHouses$fuel), ]
# head(fuel_electric)
```


```{r chunk2, echo = FALSE}
lm_base = lm(price ~ lotSize + age + livingArea + pctCollege + bedrooms + fireplaces +
           heating + bathrooms + rooms + fuel + centralAir + new, data = SaratogaHouses)
lm_own = lm(price ~ lotSize + pctCollege  + heating + bathrooms + bedrooms 
         + rooms + fuel + centralAir + new + landValue + new*lotSize
         + centralAir*heating + pctCollege*age + landValue*fuel + heating*bedrooms
         , data = SaratogaHouses)
summary(lm_base)
summary(lm_own)
```


```{r chunk3, echo = FALSE, include = FALSE}
N = nrow(SaratogaHouses)
train = round(0.8*N)
test = (N-train)

rmse = function(y, yhat) {
  sqrt( mean( (y - yhat)^2 ) )
}
```


```{r chunk4, echo = FALSE, include = FALSE}
rmse1 <- NULL
rmse2 <- NULL
for (i in seq(1:200)){
  #Picking data up for training and testing
  train_cases = sample.int(N, train, replace=FALSE)
  test_cases = setdiff(1:N, train_cases)
  
  #Define the train-test data sets (for all X's and Y)
  saratoga_train = SaratogaHouses[train_cases,]
  saratoga_test = SaratogaHouses[test_cases,]
#Training
#Base model
lm1 = lm(price ~ lotSize + age + livingArea + pctCollege + bedrooms + fireplaces +
           heating + bathrooms + rooms + fuel + centralAir + new , data=saratoga_train)
#Hand-built Model
lm2 = lm(price ~ lotSize + pctCollege  + heating  + bathrooms + bedrooms 
         + rooms + fuel + centralAir + new + landValue + new*lotSize
         + centralAir*heating + pctCollege*age + landValue*fuel + heating*bedrooms
         , data=saratoga_train)

#Testing 
yhat_test1 = predict(lm1, saratoga_test)
yhat_test2 = predict(lm2, saratoga_test)

#Run it on the actual and the predicted values
rmse1[i]= rmse(saratoga_test$price, yhat_test1)
rmse2[i]= rmse(saratoga_test$price, yhat_test2)
}
```

#Averaging the out-of-sample rmse
```{r chunk5, echo = FALSE}
 mean(rmse1)
 mean(rmse2)
```

#KNN (Non-parametric model)
```{r chunk6, echo = FALSE, include = FALSE}
knn_model = do(100)*{
N = nrow(SaratogaHouses)
train = round(0.8*N)
test = (N-train)

train_cases = sample.int(N, train, replace=FALSE)
test_cases = setdiff(1:N, train_cases)
  
saratoga_train = SaratogaHouses[train_cases,]
saratoga_test = SaratogaHouses[test_cases,]

Xtrain = model.matrix(~ lotSize + pctCollege  + heating  + bathrooms + bedrooms 
                      + rooms + fuel + centralAir + new + landValue - 1, data=saratoga_train)
Xtest = model.matrix(~ lotSize + pctCollege  + heating  + bathrooms + bedrooms 
                     + rooms + fuel + centralAir + new + landValue - 1, data=saratoga_test)
Ytrain = saratoga_train$price
Ytest = saratoga_test$price

#Scaling the features (Standardization)
scale_train = apply(Xtrain, 2, sd) 
Xtilde_train = scale(Xtrain, scale = scale_train)
Xtilde_test = scale(Xtest, scale = scale_train) 

#The for loop 
library(foreach)
k_grid = seq(2,100)
rmse_grid = foreach(K = k_grid, .combine='c') %do% {
  knn_model = knn.reg(Xtilde_train, Xtilde_test, Ytrain, k=K)
  rmse(Ytest, knn_model$pred)
}
}
 knn_model_mean = colMeans(knn_model)
```
 
#Plot KNN vs RMSE
```{r chunk7, echo = FALSE}
#Plotting 
plot(k_grid, knn_model_mean)
abline(h=rmse(Ytest, yhat_test2))
```



#Summary: 
I found that variables telling the same thing could be eliminated. For example, if you just want to know if the house is new or not then just knowing the age is sufficient and makes the binary variable 'new' unecessary to know. Having said this I couldn't eliminate rooms because it seems like bathrooms and bedrooms are not the only type of rooms looked out for when deciding the former's number in a house. I tested if newer houses affect lot size and found that they increase pricing. I also tested if heating (hot air) affected the central air enough to affect pricing and found that it did. It looks like age of the house is correlated with the percentage of college graduates residing in the neighbourhood and the higher either of those numbers, the higher would the price be. The fuel availability is correlated with the land value and that in turn affects the price as well. 